### 2016-11-09 paper/stub.tex
work on finishing results; filled in intro/conclusion

\begin{deluxetable*}{cccccc}
\tablecaption{Catalog of candidate co-moving pairs \label{tab:catalog}}
\tablehead{\colhead{star1 source id} & \colhead{star2 source id} & \colhead{separation} & \colhead{$\ln \mathcal{L}_1 /\mathcal{L}_2$} & \colhead{ID$_\mathrm{network}$} & \colhead{$N_\mathrm{network}$}\\ \colhead{ } & \colhead{ } & \colhead{$\mathrm{pc}$} & \colhead{ } & \colhead{ } & \colhead{ }}
\startdata
249282341403694592 & 441371523899475840 & 4.9 & 7.0 & 1 & 125 \\
249087281167662464 & 441901694662492928 & 8.3 & 7.1 & 1 & 125 \\
64933755122821120 & 66786500935624320 & 7.8 & 7.5 & 0 & 151 \\
436536249718223744 & 441356921010671232 & 6.5 & 6.5 & 1 & 125 \\
3953625835302703488 & 4008706729289355520 & 6.1 & 6.3 & 8 & 47 \\
63730305286697600 & 65188085906203520 & 7.0 & 6.6 & 0 & 151 \\
1873311936758998016 & 1873312074197947392 & 7.5 & 7.2 & 4013 & 2 \\
5814505765886192512 & 5814997419380722944 & 6.1 & 6.2 & 14 & 20 \\
350903157411208832 & 352510643410053632 & 9.9 & 6.8 & 565 & 2 \\
64114241002810496 & 67618281484716544 & 7.2 & 6.1 & 0 & 151
\enddata
\tablecomments{Table 1 is published in its entirety in the machine-readable format. A portion is shown here for guidance regarding its form and content.}
\end{deluxetable*}