### 2017-02-14 paper/table_meta.tex
new catalog

\begin{table}[htb]
\centering
\caption{Candidate co-moving pairs catalog description} \label{tab:table_meta}
\begin{tabular}{l|l|l}
\hline\hline
Column Name    & Unit & Description                         \\
\hline
\multicolumn{3}{c}{Table: Star (10,606 rows)}                                         \\
\hline
row id         &      & zero-based row index                      \\
TGAS source id &      & unique source id from \tgas               \\
RA             & deg  & right ascension from \tgas                \\
DEC            & deg  & declination from \tgas                    \\
parallax       & mas  & unique source id from \tgas               \\
distance       & pc   & unique source id from \tgas               \\
$G$            & mag  & \gaia\ $G$-band magnitudes                \\
$J$            & mag  & \tmass\ $J$-band magnitudes               \\
$H$            & mag  & \tmass\ $H$-band magnitudes               \\
RAVE OBS ID    &      & unique id of the \rave\ match             \\
RV             & \kms & radial velocity from \rave\               \\
eRV            & \kms & uncertainty of radial velocity from \rave \\
group id       &      & Id of the group this star belongs to      \\
\hline
\multicolumn{3}{c}{Table: Pair (13,058 rows)}                                            \\
\hline
angsep                           & arcmin & angular separation                  \\
separation                       & pc     & physical separation                 \\
$\ln\mathcal{L}_1/\mathcal{L}_2$ &        & likelihood ratio                    \\
group id                         &        & id of the group the pair belongs to \\
\hline
\multicolumn{3}{c}{Table: Group (4,555 rows)}                                         \\
\hline
id     &        & unique group id                     \\
size   &        & number of stars in a group          \\
       &        &                                     \\
\hline\hline
\end{tabular}
\end{table}